---
title: 'Feature selection'
author: 'Koen Rademaker'
date: '10 April 2020'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```

The dataset consists of 100 array Comparative Genomic Hybridization (aCGH) samples measured over 2834 genomic regions.

### Load required packages, set global variables
```{r}
library(ggplot2)
library(ggpubr)
library(tidyr)
library(caret)
library(plyr)
library(dplyr)
root_dir = getwd()
```

### Custom functions
```{r}
attach_clinical_labels <- function(df_call, df_clinical){
  tmp_df <- merge(df_call, df_clinical, by = 'row.names')
  row.names(tmp_df) <- tmp_df$Row.names
  tmp_df$Row.names <- NULL
  return(tmp_df)
}

detach_clinical_labels <- function(df){
  df$Subgroup <- NULL
  return(df)
}
```


### Loading the dataset
```{r}
########## aCGH data
# load called log2-ratios of copy numbers between diseased and healthy tissues
train_call <- read.delim(file.path(root_dir, 'data', 'train_call.tsv'), header = TRUE, sep = '\t')
# add ID for prospective features (genomic regions)
train_call$ID <- seq.int(nrow(train_call))
# store data on genomic regions separately
genomic_regions <- train_call[c('ID', 'Chromosome', 'Start', 'End', 'Nclone')]
# convert aCGH to purely numerical data, transpose to set genomic regions as features
train_call <- subset(train_call, select = -c(Chromosome, Start, End, Nclone))
train_call <- as.data.frame(t(subset(train_call, select = -c(ID))))

########## clinical data
# load and format clinical data
train_clinical <- read.delim(file.path(root_dir, 'data', 'train_clinical.tsv'), header = TRUE, sep = '\t')
row.names(train_clinical) <- train_clinical$Sample
train_clinical <- train_clinical[,-(1:1), drop=FALSE]
# display the the distribution of clinical subgroups
summary(train_clinical)
```


### PCA
```{r}
########## Principal Component Analysis
# merge data with clinical labels for visualization later on
train_call.pca_data <- merge(train_clinical, train_call, by = 'row.names')
# remove row names
train_call.pca_data$Row.names <- NULL
# run PCA with scaling applied
train_call.pca <- prcomp(train_call.pca_data[,c(2:length(train_call.pca_data))], scale. = TRUE)
# format subgroup-annotated data for PC 1 & 2
train_call.pc_1_2_annotated <- data.frame('Subgroup' = train_call.pca_data$Subgroup, train_call.pca$x[, 1:2])
# extract variances per PC
train_call.pca.variances <- round(train_call.pca$sdev^2/sum(train_call.pca$sdev^2) * 100, digits = 2)
# generate plot of samples for PC 1 & 2, coloured according to subgroup
pca_plot <- ggplot(data = train_call.pc_1_2_annotated) + geom_point(aes(x = PC1, y = PC2, col = Subgroup)) + labs(x = paste('PC1 (', train_call.pca.variances[1], '%)', sep=''), y = paste('PC2 (', train_call.pca.variances[2], '%)', sep = '')) + theme_minimal()
# print plot
print(pca_plot)
# print a summary of PCA and explained (cumulative) variance per PC
summary(train_call.pca)
```


### Simplistic statistical testing
```{r}
# attach clinical labels to samples and create continuous data frame with copy numbers
train_call.gathered <- attach_clinical_labels(train_call, train_clinical)
train_call.gathered <- train_call.gathered %>% gather(feature, copy_number, 1 : (length(train_call.gathered) - 1))

# create individual histograms of copy number distributions per subgroup
hist_her2 <- ggplot(data = train_call.gathered[train_call.gathered$Subgroup=='HER2+',], aes(x = copy_number)) + geom_histogram(binwidth = 0.5, fill = '#F8766D') + labs(x = 'Copy number', y = 'Count', title = 'HER2+') + ylim(0, 65000) + theme_classic()
hist_hr <- ggplot(data = train_call.gathered[train_call.gathered$Subgroup=='HR+',], aes(x = copy_number)) + geom_histogram(binwidth = 0.5, fill = '#00BA38') + labs(x = 'Copy number', y = 'Count', title = 'HR+') + ylim(0, 65000) + theme_classic()
hist_triple_neg <- ggplot(data = train_call.gathered[train_call.gathered$Subgroup=='Triple Neg',], aes(x = copy_number)) + geom_histogram(binwidth = 0.5, fill = '#619CFF') + labs(x = 'Copy number', y = 'Count', title = 'Triple negative') + ylim(0, 65000) + theme_classic()

# merge individual histograms into combined figure to obtain insights in copy number events per subgroup
copy_number_events_plot <- ggarrange(hist_her2, hist_hr, hist_triple_neg,
                                     ncol = 3, nrow = 1)
copy_number_events_plot
```

```{r}
# extract clinical outcomes of samples
train_call <- attach_clinical_labels(train_call, train_clinical)
clinical_outcomes <- as.factor(train_call$Subgroup)
train_call <- detach_clinical_labels(train_call)
# assess variance importance with the calculaton of filter-based variable importance
filter_variance_importances <- filterVarImp(train_call, clinical_outcomes, nonpara = FALSE)
# calculate average importance of features over all three clinical subgroups
filter_variance_importances$mean <- apply(filter_variance_importances, 1, mean)
# sort features by importance
filter_variance_importances <- filter_variance_importances[order(-filter_variance_importances$mean),]

# plot copy number patterns of clinical subgroups for top 30 important features
train_call <- attach_clinical_labels(train_call, train_clinical)
for (feature in row.names(filter_variance_importances)[1:30]) {
  print(ggplot(data = train_call, aes_string(x = as.name(feature), y = 'Subgroup', colour = 'Subgroup')) + geom_point() + labs(x = paste('Feature', feature), y = 'Clinical subgroup') + theme_classic())
}
train_call <- detach_clinical_labels(train_call)
```

Some concluding remarks on these features, the following should be considered as putative clinically discriminative features: 2185, 2222, 855, 853, 850, 1657, 856, 1673, 674, 852, 2057, 849. This consideration arises from the phenomena where at least one subgroup displays a unique pattern, e.g. only HER2+ has 2 or HR+ has no loss.

```{r}
# apply Pearson's chi-squared test to each feature
for (feature in colnames(train_call)){
  # extract feature data
  feature_data <- train_call[feature]
  # attach clinical labels
  feature_data <- attach_clinical_labels(feature_data, train_clinical)
  # tally counts of all copy number-subgroup combinations
  feature_tally <- group_by(feature_data, !! sym(feature), Subgroup) %>% tally(name = 'count')
  print(feature)
  print(feature_tally)
      #feature_contingency_table <- xtabs(count~Subgroup+feature, data = feature_tally)
}

tmp <- train_call[2185]
tmp <- attach_clinical_labels(tmp, train_clinical)
tmp_cont <- group_by(tmp, `2185`, Subgroup) %>% tally(name = 'count')
tmp_tabs <- xtabs(count~Subgroup+`2185`, data = tmp_cont)
chisq.test(tmp_tabs, simulate.p.value = TRUE)
# for each feature f
  # construct a contingency table according to:
  # tmp %>% group_by(`2185`, Subgroup) %>% tally()
  #             copy number
  #   subgroup  -1   0   1   2
  # HER2+       x    x   x   x  
  # HR+         x    x   x   x
  # TN          x    x   x   x
  #
  # perform Pearson chi-squared test
  #
  # store htest objects as variables in a chi-squared results object
  # 
  # construct table with features as rows, and columns P-value, sign_0.05 and sign_bonf
  #
  # test significance both with default 5% cut-off value and Bonferroni correction of 0.05 / 2834.
```


### Results