---
title: "CATS-Cross-Validation-Boruta-KNN"
author: "Joris Visser"
date: "12-5-2020"
output: html_document
---

#The dataset consist of an aCGH (array comparative genomic hybridization) with 100 samples of three different breast cancer subtypes. The chromosomal copy number abberations have been measured by comparison to regular DNA and from this the specific type of 

```{r}
#### Install packages
library(tidyr)
library(dplyr)

root_dir = getwd()
#### Loading the aCGH data
#Load all the data (log2-scaled of copy number aberations between breast cancer subtypes)
train_call <- read.delim(file.path(root_dir,'data', 'Train_call.txt'), header = TRUE, sep='\t')

train_call$ID <- seq.int(nrow(train_call))
genomic_regions <- train_call[c('ID','Chromosome','Start','End','Nclone')]

train_call <- subset(train_call, select = -c(Chromosome,Start,End,Nclone))
train_call <- as.data.frame(t(subset(train_call, select = -c(ID))))

#### Clinical data
train_clinical <- read.delim(file.path(root_dir, 'data', "Train_clinical.txt"), header=TRUE, sep="\t")
row.names(train_clinical) <- train_clinical$Sample
train_clinical <- train_clinical[,-(1:1), drop=FALSE]

#### Merged clinical & array data
train_data <- merge(train_call, train_clinical, by='row.names')
rownames(train_data) <- train_data$Row.names
train_data$Row.names <- NULL
train_data <- train_data %>% select(Subgroup, everything())

```

#For this part I will provide a cross-validation scheme. We will first split the data into a test and a training set. The training set will be used to do feature selection and create the model with various classification algorithms.

#For the scheme, we will split the data set into a test (for validation of final predictor) and training set (from which we select features and build the classification model). This outer loop will be 10-fold cross-validated


```{r}
############## Cross-Validation & Model
### Load packages
library('class')
library('gmodels')
library('caret')
library('Boruta')
library('randomForest')

### Parameters
#Outer loop CV
cv_fold_outer <- 10
training_size_outer <- 0.7
validation_size_outer <- 1 - training_size_outer
#Inner loop CV
cv_fold_inner <- 10
feature_size <- 0.7
#Variable importance
threshold.VarImp <- 0.7
#Statistical-testing
default_sign_threshold <- 0.05

### Functions
# Split the input data
split <- function(train, trainingsize){
  trainIndex <- createDataPartition(train$Subgroup, p = trainingsize)
  #Save the sampleIDs with index from partition & filter out the corresponding samples for training
  resample.train <- rownames(train)[trainIndex$Resample1] 
  train_set <- filter(train, rownames(train) %in% resample.train) 
  rownames(train_set) <- resample.train 
  #Save the sampleIDs with index from partition & filter out the corresponding samples for validation
  resample.validation <- rownames(train)[-trainIndex$Resample1] 
  validation_set <- filter(train, rownames(train) %in% resample.validation) 
  rownames(validation_set) <- resample.validation 
  return(list(train_set, validation_set))
}

#Feature selection with VarImp
features.VarImp <- function(data, clinical_outcome){
  filter_variance_importances <- filterVarImp(data, clinical_outcome, nonpara = TRUE)
  filter_variance_importances$mean <- apply(filter_variance_importances, 1, mean)
  filter_variance_importances <- filter_variance_importances[order(-filter_variance_importances$mean),]
  filter_variance_importances$features <- rownames(filter_variance_importances)
  features <- as.numeric((filter(filter_variance_importances, filter_variance_importances$mean > threshold.VarImp))$features)
  return(features)
}

#Feature selection with Chi-Squared test
features.chi_squared <- function(data, clinical_outcome){
  chi_squared <- list()
  chi_squared_simulated <- list()
  for (feature in colnames(data)){
    # Extract feature data
    feature_data <- data[feature]
    colnames(feature_data) <- 'copy_number'
    # Attach clinical labels
    feature_data <- attach_clinical_labels(feature_data, clinical_outcome)
    # tally counts of all copy number-subgroup combinations
    feature_tally <- group_by(feature_data, copy_number, Subgroup) %>% tally(name = 'count')
    # Construct contigency table
    feature_contingency_table <- xtabs(count~Subgroup+copy_number, data = feature_tally)
    # Perform Pearson's chi-squared test
    chi_squared[[feature]] <- chisq.test(feature_contingency_table, simulate.p.value = TRUE)
  }
  # Set bonferroni significance threshold
  bonferroni_sign_threshold <- default_sign_threshold / length(data)
  # Perform chi-squared statistical test with Monte Carlo simulations
  chi_squared.results <- data.frame(P_value = sapply(chi_squared, function(x) x$p.value), P_value_bonferroni = sapply(chi_squared, function(x) x$p.value / length(data)))
  chi_squared.significance <- data.frame(default = sapply(chi_squared.results$P_value, function(x) x < default_sign_threshold), bonferroni = sapply(chi_squared.results$P_value_bonferroni, function(x) x < bonferroni_sign_threshold))
  # Selecting the features below the bonferroni significance threshold
  chi_squared.significance$features <- rownames(chi_squared.significance)
  features <- as.numeric((filter(chi_squared.significance, chi_squared.significance$bonferroni)$features))
  return (features)
}

#Feature selection with Boruta
features.boruta <- function(z){
  print('Running Boruta..')
  boruta <- Boruta(Subgroup~ ., data = z, doTrace = 1, maxRuns = 1000)
  Tboruta <- TentativeRoughFix(boruta)
  #k <- z[Tboruta$finalDecision == "Confirmed"]
  #k$Subgroup <- z$Subgroup
  #k$Tboruta <- Tboruta
  return(Tboruta)
}
#### Prediction function for KNN
pred_knn <- function(training_set, test_set, k = ""){
  ctrl <- trainControl(method = "cv",  # cross-validation
                       number = 3,  # 10 folds
                       classProbs = FALSE , # report class probability
  )
  
  #If k is not given, do hyperparameter tuning
  if (k == ""){
    knnGrid <- expand.grid(k = c(1:10))
  } else {
    knnGrid <- expand.grid(k = k) #if k is given, use that hyperparameter
  }
  
  #Only use the data, not the subtypes
  training_set.data <- data.frame(subset(training_set, select =   -c(Subgroup)))
  test_set.data <- data.frame(subset(test_set, select = -c(Subgroup)))
  
  model.knn <- train(training_set.data, training_set$Subgroup,
                    method = 'knn',
                    tuneGrid = knnGrid
  )
  print(model.knn)
  
  #Make predictions
  predictions.knn <- predict(model.knn, newdata = test_set.data) 
    
  #Return the predictions, testset and the KNN model
  return(list('predictions'=predictions.knn, 'test'=test_set, 'model'=model.knn))
  
  #Select columns with selected features only
  #selected.train <- select(training, features)
  #selected.test <- select(test, features)
  #cl <- training$Subgroup
  
  #Make predictions
  #prediction <- knn(selected.train[,-c(1)], selected.test[,-c(1)], cl, k=k_number, prob=TRUE)
  #return(list('predictions'=prediction, 'test'=selected.test))
}

### Prediction function for Neural Networks
pred_nn <- function(training_set, test_set, size = "", decay = ""){
  #CV
  ctrl <- trainControl(method = "cv",  # cross-validation
                       number = 3,  # 10 folds
                       classProbs = FALSE , # report class probability
  )
  
  if (decay == "" | size == ""){
    nnetGrid <- expand.grid(decay = c(0.1, 0.01, 0.001),  size = c(5:7))
  } else {
    nnetGrid <- expand.grid(decay = c(decay), size = c(size))
  }
  maxSize <- max(nnetGrid$size)
  training_set.data <- data.frame(subset(training_set, select =   -c(Subgroup)))
  test_set.data <- data.frame(subset(test_set, select = -c(Subgroup)))
  
  numWts <- 1*(maxSize * (length(training_set.data) + 1) + maxSize + 1)
  print("NN working")
  model.nn <- train(training_set.data, training_set$Subgroup,
                    method = "nnet", # train neural network using `nnet`
                    tuneGrid = nnetGrid, # tuning grid
                    #preProc = c("scale"), # standardize data
                    trace = FALSE,  # hide the training trace
                    #trControl = ctrl,
                    MaxNWts = numWts*10,
                    maxit = 100 # maximum iteration
  )
  print(model.nn)
  predictions.nn <- predict(model.nn, newdata = test_set.data) 
    
  return(list('predictions'=predictions.nn, 'test'=test_set, 'model'=model.nn))
}
#####Start program
final_accuracy <- list()
final_performance_list = list(sensitivitiy=list(), specificity=list(), recall=list(), f1=list())

### Most outer loop - divide the training & validation set
#For a 10-fold CV
for (k in 1:cv_fold_outer) {
  #Set new seed for every iteration
  set.seed(k)
  #Use partition function to split data set into a training & validation set
  training_test_set <- split(train_data, training_size_outer)
  training_set <- training_test_set[[1]]
  test_set <- training_test_set[[2]]

  #Set list to store accuracies, features and other performance metrics
  accuracy_list = list()
  feature_list = list(feature=list())
  performance_list = list(sensitvitiy=list(), specificity=list(), recall=list(), f1=list())
  
  #list for hyperparameters
  hyperparameters_list = list()
  
  ####Inner loop - training
  for (i in 1:cv_fold_inner) {
    #Check progress
    print(k)
    print(i)
    
    #Set new seed for every iteration
    set.seed(69+i)
    
    #New training sets for the feature selection from the training set
    training_test_set.feature <- split(training_set, feature_size)
    training_set.feature <- training_test_set.feature[[1]]
    test_set.feature <- training_test_set.feature[[2]]
    
    #Boruta - feature selection
    Tboruta <- features.boruta(training_set.feature)
    training_set.feature.data <- training_set.feature %>% select(-Subgroup)
    feature_selection.boruta_train <- training_set.feature.data[Tboruta$finalDecision == "Confirmed"]
    features.boruta_train <- colnames(feature_selection.boruta_train)
    feature_selection.boruta_train$Subgroup <- training_set.feature$Subgroup
    test_set.feature.data <- test_set.feature %>% select(-Subgroup)
    feature_selection.boruta_test <- test_set.feature.data[Tboruta$finalDecision == "Confirmed"]
    feature_selection.boruta_test$Subgroup <- test_set.feature$Subgroup
    
    ###Model
    #NNet
    #model.inner = pred_nn(feature_selection.boruta_train, feature_selection.boruta_test)
    #KNN
    model.inner = pred_knn(feature_selection.boruta_train, feature_selection.boruta_test)
    print("NN finished")
    
    #Evaluate overall accuracy for feature selection
    score.inner = confusionMatrix(test_set.feature$Subgroup, model.inner$predictions)
    
    #Evaluate performance per class for feature selection
    performance.df <- as.data.frame(score.inner[["byClass"]])
    performance_list$sensitivity[[i]] <- performance.df$Sensitivity
    performance_list$specificity[[i]] <- performance.df$Specificity
    performance_list$recall[[i]] <- performance.df$Recall
    performance_list$f1[[i]] <- performance.df$F1
    #Append the accuracy and corresponding features to lists
    accuracy_list[i] <- score.inner[["overall"]][["Accuracy"]]
    feature_list$feature[[i]] <- features.boruta_train
    
    ###Store hyperparameters in list
    #NNet
    #hyperparameters_list$size[[i]] <- model$model$bestTune$size
    #hyperparameters_list$decay[[i]] <- model$model$bestTune$decay
    
    #KNN
    hyperparameters_list$k[[i]] <- model.inner$model$bestTune$k
  }
  avgs_vector <- unlist(accuracy_list)
  top_index <- which.max(avgs_vector)
  top_features <- feature_list$feature[[top_index]]
  
  training.features <- training_set[,top_features]
  test.features <- test_set[,top_features]
  
  training_set.features <- cbind(training_set$Subgroup, training.features)
  test_set.features <- cbind(test_set$Subgroup, test.features)
  colnames(training_set.features)[1] <- 'Subgroup'
  colnames(test_set.features)[1] <- 'Subgroup'
  
  ###NNet
  #best_size <- unlist(hyperparameters_list$size)[top_index]
  #best_decay <- unlist(hyperparameters_list$decay)[top_index]
  #model.outer = pred_nn(training_set.features, test_set.features, best_size, best_decay)
  ##Save the models in a list
  #saveRDS(model.outer$model, paste(root_dir, '/NNetmodel_', k, '.rds', sep=''))
  
  ###KNN
  best_k <- unlist(hyperparameters_list$k)[top_index]
  model.outer = pred_knn(training_set.features, test_set.features, best_k)
  ##Save the models in a list
  saveRDS(model.outer$model, paste(root_dir, '/KNNmodel_', k, '.rds', sep=''))

  score.outer = confusionMatrix(test_set$Subgroup, model.outer$predictions)
  #Accuracy
  acc = score.outer[["overall"]][["Accuracy"]]
  final_accuracy[k] <- acc
  
  #Performance metrics - Precision, Recall, Sensitvity, Specificity per class
  final_performance.df <- as.data.frame(score.outer[["byClass"]])
  final_performance_list$sensitivity[[k]] <- final_performance.df$Sensitivity
  final_performance_list$specificity[[k]] <- final_performance.df$Specificity
  final_performance_list$recall[[k]] <- final_performance.df$Recall
  final_performance_list$f1[[k]] <- final_performance.df$F1
}

best_model_number <- which.max(unlist(final_accuracy))
mean(unlist(final_accuracy), na.rm = TRUE)

```